---
title: "Project_1_Adriana"
author: "Adriana Medina"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction
In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:
 Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents


## Load Packages
```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
library(stringr)
library(tibble)
library(knitr)
```

## Upload chess tournament data
```{r}
raw_data_path<-("https://raw.githubusercontent.com/amedina613/Project-1-Adriana/main/chess_tour.txt")

raw_data <- readLines(raw_data_path)
print(raw_data[1:6])
```
## Clean data, get rid of the dashes

```{r}
raw_data = str_replace_all(raw_data,"-","") 
print(raw_data [1:6])
```
## We need to convert these lines into a dataset
```{r } 
r_data_set = read.delim(textConnection(raw_data),header = F, sep = "|", stringsAsFactors = F)
```

## The last column is blank. Remove last column

```{r}
r_data_set = r_data_set[,-11]
```

# Merge header rows
```{r}
header = paste(trimws(as.character(r_data_set[1,])), trimws(as.character(r_data_set[2,])))

header = make.names(str_sub(header, 1,11))
```

```{r}
r_data_set = as_tibble(r_data_set[-1:-2,])
names(r_data_set) = header
```

# seq: This is a function in R used to generate sequences of numbers.
+ 1: This is the starting value of the sequence. In this case, it starts at 1.
+ nrow(r_data_set): This returns the number of rows in the data frame r_data_set. 
+ nrow() is a function in R that provides the number of rows in a matrix or data frame.
+ 2: This is the increment or step size. In this case, it's 2, so the sequence will skip every other number.
```{r} 
player_id = seq(1,nrow(r_data_set),2)
states_ranks_id = seq(2,nrow(r_data_set),2)
```

```{r}
chess_players = r_data_set[player_id,]
state_ranks = select(r_data_set[states_ranks_id,], State = Pair.Num, Pre.Rating = Player.Name)
```

# Use Regex to filter out the Pre-rating value we need:

+ .*: Matches any characters before "R:".
+ R:: Matches the literal "R:".
+ \\s*: Matches zero or more whitespace characters after "R:".
+ ([0-9]{3,4}): Matches and captures between three and four digits.
+ .*: Matches any characters after the digits.

```{r}
pattern <- ".*R:\\s*([0-9]{3,4}).*"

state_ranks$Pre.Rating <- gsub(pattern, "\\1", state_ranks$Pre.Rating)

```
# Combine columns
```{r}
chess_players = bind_cols(chess_players, state_ranks)

print(chess_players[1:5,])
```

# The Round Columns contain two kinds of information in them. We just need a column with the opponent number and another with the opponent's corresponding pre-rating. 
pivot_longer from tidyr allows me too transpose the Round #s and pivot the info from a wide format to a long format, meaning I am increasing the number of rows and decreasing the number of columns.

```{r}
chess_players <- chess_players %>%
  pivot_longer(cols = starts_with("Round"), names_to = "Round", values_to = "Opponent.Num")
```

# Use regex to extract the opp number from the letter
```{r}
chess_players <- chess_players %>%
  mutate(Opponent.Num = str_extract(Opponent.Num, "\\d+"))
```

# Add a column with corresponding Opponent Rating. Had trouble with this but research said to make sure the columns I am trying to work with have compatible data types.

```{r}
str(chess_players)
```
# They were not compatible so let's try to make the numbers interpretable as numbers
```{r}
chess_players = chess_players %>% 
  mutate(Total.Pts = as.numeric(Total.Pts),
         Pair.Num = parse_number(Pair.Num),
         Player.Name = trimws(Player.Name),
         Opponent.Num = as.numeric(Opponent.Num),
         Pre.Rating = as.numeric(Pre.Rating))
```

# Now let's try to add a new column with the Opponent Rating
```{r}
chess_players <- chess_players %>%
  mutate(Opponent.Rating = Pre.Rating[match(Opponent.Num, Pair.Num)])

print(chess_players[1:10,])
```
# Calculate the avg pre chess rating of opponents, there were some N/A values so remove N/A values with na.rm
```{r}
chess_players <- chess_players %>%
  group_by(Player.Name) %>%
  mutate(Avg.pcr = mean(Opponent.Rating, na.rm = TRUE))

print(chess_players[1:10,])
```

# Still need : Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

This line was entered after feedback: my player's data was repeating itself 7 times per player, representing each round. 
```{r}
round_1_data <- chess_players %>%
  filter(Round == "Round.1")
```


```{r}
columns_keep <- c("Player.Name", "State", "Total.Pts", "Pre.Rating", "Avg.pcr")
final_chess_data<- round_1_data[columns_keep]

print(final_chess_data[1:10,])
```

```{r}
write.csv(final_chess_data, file = "~/Documents/GitHub/Project-1-Adriana/final_chess_data.csv", row.names = FALSE)
```


